MSF_Groupwork-NBA_shotlog

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
debugs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
Date submitted: 16/10/2016

@@

#Report structure

@@

@@

 The data has been acquired from Kaggle website **(https://www.kaggle.com/dansbecker/nba-shot-logs)** on 30 September 2016.

@@

##Analysis

@@

This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to when playing 'away'. Esssentially, this analysis focuses on effect of the game lovation (home/away) on winning ratio.

The NULL Hypothesis: Teams playing at home is more likely to win then when playing away.

###Methods
Steps to test the hypothesis are as follows:  
    1. Calculate a total number of games, which has been won by a home or away team
    2. Compare the total win numbers to identify whether home advantage is evident using means of the variables
    3. Extend the analysis to review of the home/away wins on a team level 
    4. Plot the team level data 
    5. Discuss the findings of the analysis

```{r, include=FALSE}
#preparing a dataframe for home advantage analysis
aa <- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)

#table showing a number of games won by home and away teams in the whole season
table(aa$WINNER)

#home wins:
hh <- sum(aa$WINNER == "HOME")
#Away wins:
aw <- sum(aa$WINNER == "AWAY")

#percentage of home team winning: 
round(hh / (hh + aw), 4)
round(aw / (hh + aw), 4)

```{r, include=FALSE}
#preparing a dataframe for analysis of games when a team played at home 
aa1 <- cast(aa, HOME_TEAM ~ WINNER); names(aa1)[1] <- "TEAM"
aa1$WIN_PCT_HOME <- round(aa1$HOME / (aa1$HOME + aa1$AWAY), 2)

#preparing a dataframe for analysis of games when a team played at away 
aa2 <- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1] <- "TEAM"
aa2$WIN_PCT_AWAY <- round(aa2$AWAY / (aa2$HOME + aa2$AWAY), 2)

#merging two dataframes together to enable the effective analysis of win ratios for teams playing at home and away.
aa3 <- merge (aa1, aa2, by ="TEAM")

#ordering the dataframe by win percentage at home.
aa3 <- aa3[order(aa3$WIN_PCT_HOME, decreasing = TRUE),]

#ordering factors of the dataframe by win percentage at home.
aa3$TEAM <- factor(aa3$TEAM, levels = aa3$TEAM[order(-aa3$WIN_PCT_HOME)])

#creating a column which shows a difference between home and away winning percentage rate
aa3$WIN_PCT_DIFF<- aa3$WIN_PCT_HOME - aa3$WIN_PCT_AWAY

###Findings
Table 1: Proportion of total games split by Location win factor

X     | HOME WIN | AWAY WIN
----- | -------- | ---
Game  | 506      | 398
Perc  | 55.97%   | 44.03%

Tables 2 & 3: Winning ratio of teams playing at home vs games playing away

TEAM        | GSW | ATL | POR | MEM | SAS | CLE | HOU | LAC | OKC | DAL | NOP | WAS | TOR | CHI | MIL  
------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
Win % HOME  |   93| 88  | 81  | 77  | 75  | 72  | 72  | 72  | 71  | 68  | 68  | 68  | 66  | 59  | 59 
Win % AWAY  |   69| 71  | 54  | 67  | 50  | 53  | 61  | 57  | 42  | 61  | 40  | 43  | 59  | 66  | 45 
Win % Diff  |   24| 17  | 27  | 10  | 25  | 19  | 11  | 15  | 29  | 7   | 28  | 25  | 7   | -7  | 14

TEAM        | PHX | IND | BOS | CHA | UTA | DEN | BKN | SAC | MIA | DET | LAL | ORL | PHI | NYK |MI 
------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
Win % HOME  | 57  | 52  | 48  | 47  | 45  | 42  | 41  | 41  | 39  | 38  | 31  | 30  | 28  | 27  | 26 
Win % AWAY  | 47  | 35  | 32  | 41  | 35  | 30  | 44  | 30  | 48  | 39  | 23  | 31  | 16  | 14  | 18 
Win % Diff  | 10  | 17  | 16  | 6   | 10  | 12  |-3   | 11  |-9   |-1   |    8|   -1|   12|   13| 8

```{r}
#plot showing home and away win ratio per team
ggplot(aa3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "AWAY")) 
```

@@

###Variance test Analysis
```{r}

```
**JUSTING TO SEE IF THERE IS ANYTHING WE CAN INCLUDE HERE

###regression analysis

@@

Table 4: Regression Analysis

@@

The analysis shown that there is a clear evidence that a team playing at home is more likely to win with an average win probability of 55.97%. The same conclusion is evident from the analysis conducted on a team level where 25 teams are winning more games at home than away. The result shown that the NULL hypothesis should be rejected as a team playing at home is more likely to win then an away team. Regression analysis shown that playing at home gives teams on average additional 4.29 margin points.

The result of this research also confirms recent trend of decline in home advantage over the years: probability of a home team winning has reduced from around 65% in 1975-1992 to average of 60.3% in 1993-2011 and to around 58.5% for 2011-15. Home win ratio was 57.4% win in 2014-2015 season. The analysis demonstrated that the home win ratio went down even further to 56.0% in 2015-2016 season. Based on the literature review this trend could be potentially explained by the following factors (Economist, 2015):

@@

This area of analysis focuses on investigation of whether a number of rest days between games affect performance and more importantly winning ratio of a team.

The Null Hypothesis is: By having longer rest a team is likely to have a higher winning ratio.

###Methods
The steps to calculate this measurement are as follows:  
    1. Calculating number of rest days between games for all teams
    2. Calculating ratios for effects of rest days on win/loss
    3. Calculating ratios for effects of rest days on win/loss form home/awayCalculating ratios for effects of rest days on win/loss
    4. Discuss the findings of the data

@@

#Testing of rest data

#t.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])
#chisq.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])

#restdays2$W[restdays2$rest==0,]

@@

Table 5: Distribution of number of games per rest days

@@

Table 6: Win Rate split by number of rest days between games

  rest days| number of wins| win percentage
-|-|-
    0| 201| 47.07
    1| 496| 50.10
    2| 137| 53.94

Table 7: Win Rate split by number of rest days between games and location

  rest days| home wins percentage | away wins percentage|
-|-|-
    0|   53.54 |44.33 
    1 | 54.68 |  44.24
    2  | 59.57 | 46.90

@@

The analysis demonstrated that overall winning ratio improves with increase in  rest days: back-to-back games had 47.07% winning ratio, 1 day rest increased winning ratio to 50.10% and with 2 days rest the winning ration has increased even further to 53.94%.

By extending the analysis to the team level, the results have demonstrated similar tendency for home  games where winning ratio has increased from 53.54% to 54.68% and 59.57% respectively with 0, 1 and 2 days of rest. Somewhat interestingly the analysis of effect of rest days on winning ratio of away games demonstrated that there is hardly any difference between 0 and 1 days of rest (44.33% and 44.24%). However with 2 days of rest teams' average winning ratio have increased to 46.90% (2.66% increase). This perhaps could be explained by the fact that teams were able to minimise effect of travelling when having a bigger window between games.

Based on the above, the analysis shown that the null hypothesis should be rejected as there is a clear evidence that the higher rest days correspond to higher winning ratio

@@

To better understand effects of rest and home/away advantage the future research should focus on distances the teams have to travel between the games. For example if a team is travelling from one coast to another- it is likely to affect them more comparing to a team from New York 'travelling' to play an away gamein New Jersey.

@@

In this analysis fatigue was measured as the period of the game, as it can be assumed that the later in the game the higher the fatigue of the players.

@@

@@

```{r}
graph.acc <- ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()

@@

The results of the T-tests show a significant difference for the overall accuracy between period 1 and period 2, between period 3 and period 4 and between regular game time and overtime. This means that player have a higher accuracy in periods 1 and 3 than in the following one. As the table shows, the changes in accuracy are minor (from 46% to 44% over the regular game time) and often located on digit-level, so although the changes are significant they are negligible. After all, the Null hypothesis must be rejected as the statistical methods used show differences for accuracy in later periods of the game, but as they are negligible in the regular game time but it can be concluded that accuracy does not suffer for later game periods as long as the game ends in regular time. For games that reach overtime a highly significant drop in the accuracy of over 5% (45% to below 40%) is observale, stating that players are less likely to score in overtime periods that in the regular game time.
This means that elite basketball players cope really well with any occurance of fatigue regarding their shot accuracy on a regular basis. As soon as the game reaches an intensity which is beyond the regular level effects of fatigue can be found leading to lower shot accuracy.

@@

```{r}

@@

The shot clock variable had missing values for 5567 shots, equal to 4.34% of all shots on the dataset, limiting the meaningfulness of the findings in the accuracy per second part.

## Shooting Percentage Analysis
###Introduction

@@

  1. “Made” and “missed” in “FGM” column were replaced by 1 and 0 respectively.
  2. A new subset was created so that each row contained the player’s name, match details and shooting performance for that player.
  3. The number of attempts and field goal percentage of all players were calculated according to the shooting data. 
  4. The players with most outstanding performances were picked out and another subset was generated for these players.
  5. The correlation (p = .05) between the results of the matches and the performances of top scorers was investigated. 
  6. A logistic regression with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables. 
  7. Another logistic regression test was run in order to examine three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.

@@

### Find field goal percentage (FG percentage) for each match

@@

```
##a. Field Goal Attempts (FGA)

There was no correlation found between the result of a match and the number of attempts of the player made in that match. The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .794, failing to reach level of significance (p = .05).

##b.  Field Goal Percentage (FG%)

There was no correlation found between the result of a match and the field goal percentage of the player in that match. The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach level of significance (p = .05).

@@

```
Table xxx: The logistic regression for FGA and FG% of 15 players.

@@

Table xxx: The logistic regression for FGA and FG% of the 3 special players.
| | Estimate | Std. Error | t value | Pr(>t)   
|-|-|-|-| - 
|(Intercept)   | 0.042259 |  0.164816 |  0.256  |  0.798    
|FGA          | -0.002679 |  0.006468|  -0.414  |    0.679    
|FG.percentage|  0.012329|   0.002736|   4.506  |1.35e-05 ***

The box plots above shows the performances the three special players in terms of FGA and FG% respectively. The red (blue) boxes contain the data when the team lost (won) the match.

@@

From the correlation tests between the match results and players’ performances, it can be noticed that the increase in number of field goal attempts from top scorers is unlikely to make any impact on the results of the game. The correlation coefficient between the results of the game and FG% is sightly higher but these two factors are still not correlated since r(832) < .01. Meanwhile, the p-values of both correlation tests are too higher to reach the default level of significance (p = .05).

From the logistic regression test for FGA and FG%. It was found that FGA did not have any impact on the result of the match. FG% had significant correlation with the result of match. However, the anova test showed that the deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.

There was no correlation found between field goal attempts of top scorers and game results. A possible explanation is that it is easier for the opponents to focus on a certain player on defensive side. Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelyhood of winning a game. Meanwhile, the accuracy of these players has a stronger impact on the outcome of the game. This is due to fact that once a player is being efficient on the offensive side, that player can help the team by not only keeping shooting but also creating more opportunities for his teammates.

Based on results above, an additional research has been done regarding to the performances of three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward. These three players were chosen because their teams are more likely to take the victory when they score more in single game according to the bar plot and t-test.

For these three players, the p-values still did not reach the level of significance (p = .05) for FGA. However, all p-values are smaller than 0.05 when their FG%s were tested. Based on these findings, another logistic regression was done with these three players. It was shown that p < .001 and residual deviance from anova dropped by 4.498 which is about 12.3% when FG% was considered. In comparison, the deviance was above 200 when 15 players were tested.

Therefore, these three players are somehow different from others as they can contribute more to the win rates of their teams by their increasing shooting accuracy. There could be many reasons behind this, including their team compositions and the roles of the players. Further research has to be done to if one wants to find out the differences of these players and others.

The data set only contains 904 out of 1230 matches in the entire regular season. So there might be a chance that the highlight performances of certain players are missed out in our analysis. Meanwhile, it would be ideal if the contributions of all players can be analysed in more aspects including number of rebounds and assists.

@@

#Calculate shooting accuracy for each distance
distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
```

```{r, echo=FALSE}
#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Closest Defender Distance", limits = 0:24) +
  ylab("Shooting Accuracy (%)") +
  ggtitle("Closest Defender Distance vs Shooting Accuracy")
```

###Plot Result
  The explanation from the plot above are as follows:   
  1. Defender distance range from 0 to 2 feet, the accuracy varied between **42%** to **50%**  
  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
  7. Distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)

###Correlation Test
  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:

@@

```{r, echo=FALSE}
#calculate correlation
cor.test(dfZerotoAll$distanceZerotoAll, dfZerotoAll$accuracyDist1)
```

###Correlation Test Result  
  The correlation test result produced slightly higher number of correlation between defender distance with shooting accuracy for all shot distance attempts (0.59). If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy. However, when analysing distance ranges the result was somewhat different. For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:

```{r, include=FALSE}
#populate data from data frame
distanceZerotoEight <- c()
accuracyDist2 <- c()
for (i in 0:80){
  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
}

@@

```{r, echo=FALSE}
#calculate correlation
cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
```

@@

#Set a new data frame for the 3 players with best performances(higher points)
best_players_df <- head(pttsum2, 3)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```
##Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
##Date submitted: 16/10/2016

@@

@@

** TO GET UPDATED AS A FINAL STEP**

@@

 The data has been acquired from [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) website on 30 September 2016.

@@

#Analysis

@@

This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to when playing 'away'. Esssentially, this analysis focuses on effect of the game location (home/away) on winning ratio.

The NULL Hypothesis: Teams playing at home is more likely to win then when playing away. 
The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference

###Methods
Steps to test the hypothesis are as follows:  
    1. Calculate a total number of games, which has been won by a home or away team
    2. Compare the total win numbers to identify whether home advantage is evident using means of the variables
    3. Extend the analysis to review of the home/away wins on a team level 
    4. Plot the team level data 
    5. Perform a t-test (p .05) to examine whether there is a difference in the means between teams winning at home and teams winning away
    6. Linear regression with Final margin as dependent variable and Location as independent variable to examine whether playing at home results predicts a higher final margin

The graph below shows descriptive statistics for home/away wins:

```{r, include=FALSE}
#preparing a dataframe for home advantage analysis
advdf <- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)

#table showing a number of games won by home and away teams in the whole season
table(advdf$WINNER)

#home wins:
hh <- sum(advdf$WINNER == "HOME")
#Away wins:
aw <- sum(advdf$WINNER == "AWAY")

#percentage of home team winning: 
round(hh / (hh + aw), 4)
round(aw / (hh + aw), 4)
```

```{r, include=FALSE}
#preparing a dataframe for analysis of games when a team played at home 
advdf1 <- cast(advdf, HOME_TEAM ~ WINNER); names(advdf1)[1] <- "TEAM"
advdf1$WIN_PCT_HOME <- round(advdf1$HOME / (advdf1$HOME + advdf1$AWAY), 2)

#preparing a dataframe for analysis of games when a team played at away 
advdf2 <- cast(advdf, AWAY_TEAM ~ WINNER); names(advdf2)[1] <- "TEAM"
advdf2$WIN_PCT_AWAY <- round(advdf2$AWAY / (advdf2$HOME + advdf2$AWAY), 2)

#merging two dataframes together to enable the effective analysis of win ratios for teams playing at home and away.
advdf3 <- merge (advdf1, advdf2, by ="TEAM")

#ordering the dataframe by win percentage at home.
advdf3 <- advdf3[order(advdf3$WIN_PCT_HOME, decreasing = TRUE),]

#ordering factors of the dataframe by win percentage at home.
advdf3$TEAM <- factor(advdf3$TEAM, levels = advdf3$TEAM[order(-advdf3$WIN_PCT_HOME)])

#creating a column which shows a difference between home and away winning percentage rate
advdf3$WIN_PCT_DIFF<- advdf3$WIN_PCT_HOME - advdf3$WIN_PCT_AWAY

#t tests

t.test(advdf3$AWAY.y, advdf3$HOME.x)
```

```{r, echo=FALSE}
#plot showing home and away win ratio per team
ggplot(advdf, aes(x = advdf$WINNER)) + geom_bar(aes()) + ylab("game count") + xlab("winner")
```

###Findings
**Table 1**: Proportion of total games split by Location win factor

X     | Home Win | Away Win
----- | -------- | ---
Game  | 506      | 398
Perc  | 55.97%   | 44.03%

**Tables 2 & 3**: Winning ratio of teams playing at home vs games playing away

Team     | GSW | ATL | POR | MEM | SAS | CLE | HOU | LAC | OKC | DAL | NOP | WAS | TOR | CHI | MIL  
------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
Win % HOME  |   93| 88  | 81  | 77  | 75  | 72  | 72  | 72  | 71  | 68  | 68  | 68  | 66  | 59  | 59 
Win % AWAY  |   69| 71  | 54  | 67  | 50  | 53  | 61  | 57  | 42  | 61  | 40  | 43  | 59  | 66  | 45 
Win % Diff  |   24| 17  | 27  | 10  | 25  | 19  | 11  | 15  | 29  | 7   | 28  | 25  | 7   | -7  | 14

Team     | PHX | IND | BOS | CHA | UTA | DEN | BKN | SAC | MIA | DET | LAL | ORL | PHI | NYK |MI 
------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
Win % HOME  | 57  | 52  | 48  | 47  | 45  | 42  | 41  | 41  | 39  | 38  | 31  | 30  | 28  | 27  | 26 
Win % AWAY  | 47  | 35  | 32  | 41  | 35  | 30  | 44  | 30  | 48  | 39  | 23  | 31  | 16  | 14  | 18 
Win % Diff  | 10  | 17  | 16  | 6   | 10  | 12  |-3   | 11  |-9   |-1   |    8|   -1|   12|   13| 8

```{r, echo=FALSE}
#plot showing home and away win ratio per team
ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins at HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
```

@@

The T-test used to verify the differences in the means seen in the plot reached significance, t(58) = -2.39, p = .02.

###Regression analysis

@@

**Table 4**: Regression analysis for the effect of location and rest days on final score margin

@@

The analysis shows that there is a clear evidence that a team playing at home is more likely to win with an average win probability of 55.97%. The same conclusion is evident from the analysis conducted on a team level where 25 teams are winning more games at home than away. The results show that the NULL hypothesis should be rejected as a team playing at home is more likely to win than an away team. Linear regression analysis shows that playing at home gives teams on average additional 4.29 margin points.

The result of this research also confirms the recent trend of decline in home advantage over the years: probability of a home team winning has reduced from around 65% in 1975-1992 to average of 60.3% in 1993-2011 and to around 58.5% for 2011-14. The analysis demonstrated that the home win ratio went down even further to 56.0% in 2014-2015 season. Based on the literature review this trend could be potentially explained by the following factors (Economist, 2015):

@@

This area of analysis focuses on investigation of whether a number of rest days between games affect the performance and more importantly winning ratio of a team.

The NULL Hypothesis is: Teams' mean values of wins are not significantly different based on the number of rest days between games

The steps to calculate this measurement are as follows:  
    1. Calculating number of rest days between games for all teams
    3. Calculating ratios for effects of rest days on win/loss form home/away
    4. Perform a Chi-square test to examine whether there is a difference in the distributon for games won and games lost for the different restdays
    5. Calculating ratios for effects of rest days on win/loss form home/away
    6. Run a regression analysis
    7. Discuss the findings of the data

The table below shows descriptive statistics for Rest Days:

**Table 5**: Distribution of number of games per rest days

@@

#Testing of rest data
qqqq<-as.matrix(sapply(restdays4, as.numeric))

chisq.test(qqqq[2,], qqqq[3,])

@@

**Table 5**: Distribution of number of games per rest days

@@

**Table 6**: Win rate split by number of rest days between games

  Rest Days| Number of Wins|Win Percentage
-|-|-
    0| 201| 47.07
    1| 496| 50.10
    2| 137| 53.94

**Table 7**: Win rate split by number of rest days between games and location

  Rest Days| Home Wins Percentage | Away Wins Percentage|
-|-|-
    0|   53.54 |44.33 
    1 | 54.68 |  44.24
    2  | 59.57 | 46.90

The Pearson's Chi-Square test failed to reach the level of significance, X2 (36) = 42, p= .23.

@@

The analysis demonstrates that overall the winning ratio stays the same with increase in  rest days. However, when looking at invididual restdays the findings are different: back-to-back games have 47.07% winning ratio, 1 day rest increases winning ratio to 50.1% and with 2 days rest the winning ration increases even further to 53.94%.

By extending the analysis to the team level, the results demonstrate similar tendency for home games where winning ratio increased from 53.54% to 54.68% and 59.57% respectively with 0, 1 and 2 days of rest. Somewhat interestingly the analysis of effect of rest days on winning ratio of away games demonstrates that there is hardly any difference between 0 and 1 days of rest (44.33% and 44.24%). However with 2 days of rest teams' average winning ratio increases to 46.90% (2.66% increase). This perhaps could be explained by the fact that teams were able to minimise effect of travelling when having a bigger window between games.

Based on the above, the analysis shows that the null hypothesis should be rejected as there is a clear evidence that the higher rest days correspond to higher winning ratio.

@@

To better understand effects of rest and home/away advantage future research should focus on distances the teams have to travel between the games. For example if a team is travelling from one coast to another- it is likely to affect them more comparing to a team from New York 'travelling' to play an away game in New Jersey.

@@

In this analysis fatigue was measured as the period of the game, as it can be assumed that the later in the game the higher the fatigue of the players. The descriptive statistics of the shots per period can be seen below.

```{r, echo=FALSE}
bar.period <- ggplot(df) + geom_bar(aes(x = PERIOD))
bar.period + labs(title = "Number of shots taken per period") + scale_x_discrete(name ="Period", limits=c("1","2","3","4", "5", "6", "7")) + ylab("Number of shots")
```

@@

**Table 8**: Accuracy per period

@@

```{r, echo=FALSE}
graph.acc <- ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()

@@

The results of the T-tests show a significant difference for the overall accuracy between period 1 and period 2, between period 3 and period 4 and between regular game time and overtime. This means that player have a higher accuracy in periods 1 and 3 than in the following one. As the table 8 shows, the changes in accuracy are minor (from 46% to 44% over the regular game time) and often located on digit-level, so although the changes are significant they are negligible. After all, the Null hypothesis must be rejected as the statistical methods used show differences for accuracy in later periods of the game, but as they are negligible in the regular game time but it can be concluded that accuracy does not suffer for later game periods as long as the game ends in regular time. For games that reach overtime a highly significant drop in the accuracy of over 5% (45% to below 40%) is observale, stating that players are less likely to score in overtime periods that in the regular game time.
This means that elite basketball players cope really well with any occurance of fatigue regarding their shot accuracy on a regular basis. As soon as the game reaches an intensity which is beyond the regular level effects of fatigue can be found leading to lower shot accuracy.

@@

```{r, echo=FALSE}

@@

As there are much more shots taken in periods 1 to 4 than in 5 to 7 the differences in the accuracy have to be treat with caution.
The shot clock variable had missing values for 5567 shots, equal to 4.34% of all shots on the dataset, limiting the meaningfulness of the findings in the accuracy per second part.

## Shooting Percentage Analysis (James)

@@

  1. “Made” and “missed” in “FGM” column were replaced by 1 and 0 respectively.
  2. A new subset was created so that each row contained the player’s name, match details and shooting performance for that player.
  3. The number of attempts and field goal percentage of all players were calculated according to the shooting data. 
  4. The players with most outstanding performances were picked out and another subset was generated for these players.
  5. The correlation (p = .05) between the results of the matches and the performances of top scorers was investigated. 
  6. A logistic regression analysis with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables was conducted.

@@

@@

```
###Results

####a. Field Goal Attempts (FGA)

There was no correlation found between the outcome of the matches in this season and the number of attempts of the players made. The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .794, failing to reach the level of significance (p = .05).

####b.  Field Goal Percentage (FG%)

There was no correlation found between the outcome of the matches in this seaon and the field goal percentage of the players. The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach the level of significance (p = .05).

@@

```
**Table 9**: Result for the logistic regression analysis for FGA and FG% of 15 players

@@

**Table 10**: Result for the anova for FGA and FG% for 15 players

   | |           Df| Deviance| Resid. Df| Resid. Dev
-|-|-|-|- 
|NULL        |     |              | 777   |  188.73
|FGA         |   1  | 0.0014   |    776  |   188.73
|FG.percentage | 1 |  7.4664   |    775  |   181.26

An ad hoc data analysis was used as 3 of the 15 players stood out in particular. To examine whether their data result in a better model, another logistic regression analysis was performed with the same dependent and independent variables for three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.

```{r, echo=FALSE}
graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won of the three special players") + xlab("Player name") + ylab("FGA")
graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) 
graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won of the three special players") + xlab("Player name") + ylab("FG%")
grid.arrange(graph01, graph02, ncol = 2)
```

The box plots above shows the performances of the three special players in terms of FGA and FG% respectively.

**Table 11**: Result for the logistic regression analysis for FGA and FG% of the 3 special players

| | Estimate | Std. Error | t value | Pr(>t)   
-|-|-|-| - 
|(Intercept)   | 0.042259 |  0.164816 |  0.256  |  0.798    
|FGA          | -0.002679 |  0.006468 |  -0.414  |    0.679    
|FG.percentage |  0.012329 |   0.002736 |   4.506  | 1.35e-05 ***

**Table 12**: Result for the anova for FGA and FG% for the 3 special players

| |         Df| Deviance| Resid. Df| Resid. Dev
|-|-|-|-| - 
|NULL         |   |         |       147|     36.669
|FGA          |  1|   0.0498|       146|     36.619
|FG.percentage|  1|   4.4978|       145|     32.121

@@

From the correlation tests between the match results and players’ performances, it can be noticed that the increase in number of field goal attempts from top scorers is unlikely to make any impact on the results of the game. The p-values of both correlation tests are too high to reach the level of significance (p = .05).

From the logistic regression analysis for FGA and FG%, it was found that FGA did not have any impact on the result of the match. FG% is a significant predictor for the result of match. However, the anova showed that the residual deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.

There was no correlation found between field goal attempts of top scorers and outcome of the matches. A possible explanation is that it is easier for the opponent to focus on a certain player on defensive side and then the overall efficiency of the team will start to decrease. Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelyhood of winning a game. Meanwhile, the accuracy of these players has a stronger impact on the outcome of the game. This could be explained by that once a player is being efficient on the offensive side, that player can help the team not only by shooting but also by creating more opportunities for his teammates.

Based on the results above, an additional analysis was done regarding the performance of three particular players: Russell Westbrook, Kawhi Leonard and Gordon Hayward. These three players were chosen because their teams are more likely to take the victory when they score more in single game according to Diagram and t-test.

For these three players, the p-values still did not reach the level of significance (p = .05) for FGA. However, all p-values are smaller than 0.05 when their FG%s were tested. Based on these findings, another logistic regression was done with these three players. It was shown that p < .05 and residual deviance from anova dropped by 4.498 which is about 12.3% when FG% was considered. In comparison, the deviance was above 180 when 15 players were analysed. For these three players, their teams are more likely to win by 1.2% when their FG% increase by 1.

Therefore, the analysis on these three players has shown that there is a difference between them and the others as they can contribute more to the win rates of their teams by their increasing shooting accuracy. There could be various reasons behind this observation, such as their team compositions and the roles of these players. Further research has to be done to in order to find out why the performances of these special players are more decisive on the outcome.

###Limitations
The data set only covers 904 of 1230 matches in the regular season. So there might be a chance that the highlight performances of certain players are missed out in the analysis. Meanwhile, it would be ideal if more aspects of data, such as rebounds and assists, could be analysed.

@@

#Calculate shooting accuracy for each distance
distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)

library(ggplot2)
library(scales)
ggplot(data = df) + geom_bar(aes(x = df[, "CLOSE_DEF_DIST"])) +
  scale_x_discrete("Closest Defender Distance", breaks = c (0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24), limits = 0:24) + 
  ggtitle("Defender Distance Distribution")

```{r, echo=FALSE}
#Plotting the result
library(ggplot2)
library(scales)
ggplot(data = distance_level) +
  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
  scale_x_discrete("Closest Defender Distance", limits = 0:24) +  
  ylab("Shooting Accuracy (%)") + ggtitle("Closest Defender Distance vs Shooting Accuracy")
```

###Plot Result
  The explanation from the plot above are as follows:   
  1. Defender distance range from 0 to 2 feet, the accuracy varied between 42% to 50%  
  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from 44% to 49%  
  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from 41% to 46%
  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from 37% to 42%
  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from 37% to 54%
  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from 33% to 60%.  
  7. Distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance)

###Correlation Test
  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
  There was medium to strong positive correlation found between the defender distance and shooting accuracy. The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = .59, p < .05, as part of level of significance (p = .05).

@@

```{r, include=FALSE}
#calculate correlation
cor.test(dfZerotoAll$distanceZerotoAll, dfZerotoAll$accuracyDist1)
```

###Correlation Test Result  
  The correlation test result produced slightly higher number of correlation between defender distance with shooting accuracy for all shot distance attempts (0.59). If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy. However, when analysing distance ranges the result was somewhat different. For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:  
  There was medium negative correlation found between the defender distance and shooting accuracy in this range. The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .05, as part of level of significance (p = .05).
```{r, include=FALSE}
#populate data from data frame
distanceZerotoEight <- c()
accuracyDist2 <- c()
for (i in 0:80){
  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
}

@@

```{r, include=FALSE}
#calculate correlation
cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
```

@@

#Setting a new data frame for the 3 players with best performances(higher points)
best_players_df <- head(pttsum2, 3)
